---
title: "WorldClim in R"
output: html_notebook
---


```{r}
envs <- raster::getData(name = "worldclim", var = "bio", res = 2.5, lat = , lon = )

# extract environmental values at occ grid cells
locs.vals <- raster::extract(envs[[1]], occs[, c('longitude', 'latitude')])
# remove occs without environmental values
occs <- occs[!is.na(locs.vals), ]
```


## Rasters from files

Typically, when we work with rasters, we will be creating them from a file (instead of from arbitrary data values). We are going to start by looking at annual mean temperature across the globe from the WorldClim data you downloaded. To make a raster from a file, you first need to get the file path. Terminal makes this easy: open Terminal and open Finder. In Finder, navigate to the file "" and drag and drop it into your Terminal window. The file path should appear -- copy it.

**Try it yourself:** Use the `raster()` function to create a raster from the file of mean temperature data using the file path you just copied (in quotation marks!).

```{r}
bio1 <- 
```



Sometimes we want to get the raster value associated with a certain coordinate (this will be important for our sloths!). To accomplish this, we can use the `extract()` function:

```{r}
xy_coords <- c(0,0)
extract(raster1, xy_coords)
```
